### 2015-05-25 paper/TextS3/TextS3.tex

\documentclass{article}
\textheight24cm\textwidth16cm\topmargin-1cm
\oddsidemargin0cm\evensidemargin0cm

\usepackage{amsmath,amssymb}
% \usepackage[dvipdfmx]{graphicx}
\usepackage{graphicx}
\usepackage{subfigure}

\usepackage{url}
\usepackage{bm}
\usepackage{setspace} 
\usepackage{subfigure}

% \usepackage{algorithmic}
% \usepackage{algorithm}

\newtheorem{thm}{Theorem}
\newtheorem{dfn}[thm]{Definition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newcommand{\proof}{\noindent Proof.\ \ }
% \renewcommand{\algorithmiccomment}[1]{// #1}

%\pagestyle{empty}

\begin{document}

% \baselineskip 6.5mm

\vspace*{1.0cm}
{\LARGE \bf Derivation of EM algorithm}
\vspace*{0.25cm}

The complete log-likelihood including missing data $\{ \bm{z}_i \}$ for the proposed model is
\begin{equation*}
\sum_{i = 1}^I \sum_{j=1}^J \sum_{k=1}^K I(z_{i,j} = k) \Bigl( \sum_{l = 1}^L \log f_{k, l, x_{i,j,l}} + \log q_{i,k} \Bigr).
\end{equation*}
Here, we introduce the variable for conditional probability for $z_{i,j}$ given the parameters and the mutation features $\bm{x}_{i,j}$,
\begin{equation*}
\theta_{i,k, \bm{m}} = \Pr \bigl( z_{i,j} = k \big| \bm{x}_{i, j} = \bm{m}, \{ \bm{f}_{k, l} \}, \{  \bm{q}_i \} \bigr)
\end{equation*}
Note that this conditional probability just depends on the value of mutation feature $\bm{m} = (m_1, \cdots, m_L)$, not on the index $j$.
Then, the expected complete log-likelihood augmented by Lagrange multipliers is calculated as
\begin{equation*}
\sum_{i=1}^I \sum_{\bm{m}} g_{i, \bm{m} } \sum_{k=1}^K \theta_{i, k, \bm{m}}   \Bigl( \sum_{l = 1}^L \log f_{k, l, m_l} + \log q_{i,k} \Bigr)
+ \sum_{k=1}^K \sum_{l=1}^L \tau_{k,l} \bigl( 1 - \sum_{p=1}^{M_l} f_{k,l,p} \bigr)
+ \sum_{i=1}^I \rho_i \bigl( 1 - \sum_{k=1}^K q_{i,k} \bigr).
\end{equation*}
Differentiating it leads to following stationary equations: 
\begin{align*}
\sum_{i=1}^I \sum_{\bm{m}: m_l = p} g_{i, \bm{m}} \theta_{i,k,\bm{m}} - \tau_{k,l} f_{k,l,p} & = 0,\ (p=1, \cdots, M_l, k=1, \cdots, K,\ l=1, \cdots, L)., \\
\sum_{\bm{m}} g_{i,\bm{m}} \theta_{i,k,\bm{m}} - \rho_i q_{i, k} & = 0,\ (k = 1, \cdots, K, i = 1, \cdots, I).
\end{align*}
Then, by eliminating Lagrange multipliers, updating rules can be obtained.

\end{document}